\bf Calendar-based graphics for visualizing people's daily schedules

### 2019-03-12 main.Rmd
use base_foramt bookdown::pdf_document2

bibliography: bibliography.bib
output: rticles::asa_article
---

@@

The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped]. There have been 43 sensors installed that count pedestrians every hour across the inner-city area until the end of 2016 (Figure \ref{fig:ped-map}). The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. The faceted plots in Figure \ref{fig:time-series-plot} give an overall picture of the foot traffic at three different sensors over 2016. Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns.

@@

```{r ped-map, fig.width = 7, out.width = "0.7\\linewidth", fig.cap = "Map of the Melbourne city area with dots indicating sensor locations. These three highlighted sensors will be inspected in the paper: (1) the State Library---a public library, (2) Flagstaff Station---a train station, closed on non-work days, (3) Birrarung Marr---an outdoor park hosting many cultural and sports events."}
```

```{r time-series-plot, fig.height = 6, fig.cap = "Time series plots showing the number of pedestrians in 2016 measured at three different sensors in the city of Melbourne. Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number. A spike occurred at the State Library, caused by the annual White Night event on 20th of February. A relatively persistent pattern repeats from one week to another at Flagstaff Station. Birrarung Marr looks rather noisy and spiky, with a couple of chunks of missing records."}
```

```{r facet-time, fig.height = 6, fig.cap = "Hourly pedestrian counts for 2016 faceted by sensors and days of the week using lines. It primarily features two types of seasons---time of day and day of week---across all the sensors. Apparently other factors have influence over the number of pedestrians, which cannot be captured by the faceted plots, such as the overnight White Night traffic on Saturday at the State Library and a variety of events at Birrarung Marr."}
```

@@

The remainder of the paper is organized as follows. Section \ref{sec:algorithm} demonstrates the construction of the calendar layout in depth. Section \ref{sec:transformation} describes the algorithms of data transformation. Section \ref{sec:opt} lists and describes the options that come with the `frame_calendar` function. Section \ref{sec:variations} presents some variations of its usage. Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \ref{sec:case}. Section \ref{sec:discussion} discusses the limitations of calendar displays and possible new directions.

# Creating a calendar display
\label{sec:algorithm}

## Data transformation
\label{sec:transformation}

@@

\begin{equation}
  \begin{aligned}
  i &= \lceil (g \text{ mod } 35) / 7\rceil, \\
  j &= g \text{ mod } 7. \label{eq:grid}
  \end{aligned}
\end{equation}

Figure \ref{fig:month-diagram} illustrates this $(i,j)$ layout for a month where $k=5$.

```{r month-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "Illustration of the indexing layout for cells in a month, where $k$ is day of the week, $g$ is day of the month, $(i, j)$ indicates grid position."}
include_graphics("img/month.pdf")
```

To create the layout for a full year, $(m, n)$ denotes the position of the month arranged in the plot, where $1 \le m \le M$ and $1 \le n \le N$; $b$ denotes the small amount of white space between each month for visual separation. Figure \ref{fig:year-diagram} illustrates this layout where $M = 3$ and $N = 4$.

```{r year-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "Illustration of the indexing layout for months of one year, where $M$ and $N$ indicate number of rows and columns, $b$ is a space parameter separating cells."}
include_graphics("img/year-diagram.pdf")
```

@@

\begin{equation}
  \begin{aligned}
  x &= j + (n - 1) \times 7 + (n - 1) \times b + h, \\
  y &= i - (m - 1) \times 5 - (m - 1) \times b + c. \label{eq:final}
  \end{aligned}
\end{equation}

Note that for the vertical direction, the top left is the starting point of the grid (in Figure \ref{fig:month-diagram}) which is why subtraction is performed. Within each cell, the starting position is the bottom left.

```{r fs-2016, fig.height = 7, fig.cap = "The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs. The disparities between week day and weekend along with public holiday are immediately apparent. The arrangement of the data into a $3 \\times 4$ monthly grid represents all the traffic in 2016. Note that the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs to May and October."}
```

Figure \ref{fig:fs-2016} shows the line glyphs framed in the monthly calendar over the year 2016. This is achieved by the `frame_calendar` function, which computes the coordinates on the calendar for the input data variables. These can then be plotted using the usual **ggplot2** R package [@R-ggplot2] functions. All of the grammar of graphics can be applied.

@@

Regarding the monthly calendar, the major reference lines separate every month panel and the minor ones separate every cell, represented by the thick and thin lines in Figure \ref{fig:fs-2016}, respectively. The major reference lines are placed surrounding every month block: for each $m$, the vertical lines are determined by $\min{(x)}$ and $\max{(x)}$; for each $n$, the horizontal lines are given by $\min{(y)}$ and $\max{(y)}$. The minor reference lines are only placed on the left side of every cell: for each $i$, the vertical division is $\min{(x)}$; for each $j$, the horizontal is $\min{(y)}$.

@@

## Options
\label{sec:opt}

@@

It is assumed that the `data` is in tidy format [@wickham2014tidy], and `x`, `y` are the variables that will be mapped to the horizontal and vertical axes in each cell. For example, the `x` is the time of the day, and `y` is the count (Figure \ref{fig:fs-2016}). The `date` argument specifies the date variable used to construct the calendar layout.

@@

```{r fs-polar, fig.height = 7, fig.cap = "Figure \\ref{fig:fs-2016} in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible.", eval = FALSE}
```

@@

With local scaling, the overall variation gives way to the individual shape. Figure \ref{fig:fs-free} shows the same data as Figure \ref{fig:fs-2016} scaled locally using `scale = "free"`. The daily trends are magnified.

```{r fs-free, fig.height = 7, fig.cap = "Line glyphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled over all the days. The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days."}
```

@@

By default, grids are laid out horizontally. This can be transposed by setting the `dir` parameter to `"v"`, in which case $i$ and $j$ are swapped in Equation \ref{eq:grid}. This can be useful for creating  calendar layouts for countries where vertical layout is the convention.

@@

## Variations
\label{sec:variations}

@@

Plots can be layered. The comparison of sensors can be done by overlaying the values for each (Figure \ref{fig:overlay}). Differences between the pedestrian patterns at these sensors can be seen. Flagstaff Station exhibits strong commuters patterns, with fewer pedestrian counts during the weekends and public holidays. This suggests that Flagstaff Station has limited functionality on non-work days. From Figure \ref{fig:overlay} it can be seen that Birrarung Marr has a distinct temporal pattern from the other two all year round. The nighttime events, such as White Night, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to the State Library and Birrarung Marr.

```{r overlay, fig.height = 7, fig.cap = "Overlaying line graphs of the three sensors in the monthly calendar. Three sensors demonstrate very different traffic patterns. Birrarung Marr tends to attract many pedestrians for special events held on weekends, contrasting to the bimodal commuting traffic at Flagstaff Station."}
```

To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors. Figure \ref{fig:facet} presents the idea of faceting calendar plots. This allows comparing the overall structure between sensors, while emphasizing individual sensor variation. In particular, it can be immediately learned that Birrarung Marr was busy and packed, for example during the Australian Open, a major international tennis tournament, in the last two weeks of January. This is concealed in the conventional graphics.

```{r facet, fig.height = 10.2, fig.cap = "Line charts, embedded in the $6 \\times 2$ monthly calendar, colored and faceted by the 3 sensors. The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors."}
```

@@

The `frame_calendar` function is not constrained to line plots. The full range of plotting capabilities in **ggplot2** is essentially available. Figure \ref{fig:scatterplot} shows a lag scatterplot at Flagstaff Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument. This figure is organized in the daily calendar layout. Figure \ref{fig:scatterplot} indicates two primary patterns, strong autocorrelation on weekends, and weaker autocorrelation on work days. At the higher counts, on week days, the next hour sees possibly substantial increase or decrease in counts, essentially revealing a bimodal distribution of consecutive counts, as supported by Figure \ref{fig:fs-2016}.

```{r scatterplot, fig.height = 5, fig.cap = "Lag scatterplot in the daily calendar layout. Each hour's count is plotted against previous hour's count at Flagstaff Station to demonstrate the autocorrelation at lag 1. The correlation between them is more consistent on non-work days than work days."}
```

The algorithm can also produce more complicated plots, such as boxplots. Figure \ref{fig:boxplot} uses a loess smooth line [@cleveland1979loess] superimposed on side-by-side boxplots. It shows the distribution of hourly counts across all 43 sensors during December. The last week of December is the holiday season: people are off work on the day before Christmas (December 24), go shopping on the Boxing day (December 26), and stay out for the fireworks on New Year's Eve.

@@

```{r chn}
```

```{r chn-embedded, fig.cap = "The same plot as Figure \\ref{fig:boxplot}, but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English."}
include_graphics("figure/chn-1.pdf")
```

@@

# Case study
\label{sec:case}

@@

Figure \ref{fig:dow} shows the energy use across days of week in the form of letter value plots [@hofmann2017letter]. Letter value plots are a variant of boxplots for large data, with other quantiles represented by boxes. Letters indicate the fraction of the data divisions, for example, F indicates fourths or quartiles, and the two outer ends of the box are the 25th and 75th percentile, the same traditional ends of the box as a boxplot. The letter E indicates eighths, with box ends being 12.5th and 87.5th percentiles of the data. These additional boxes replace the whiskers in a traditional boxplot. The letter value plots for the households, show a line indicating the median (M) and the innermost boxes corresponding to the fourth (F) and the eighth (E) quantile divisions. Inspecting the medians across households tells us that household 3, a family size of one couple and two kids, uses more energy over the week days, than other households. The relatively larger boxes for household 2 indicates greater variability in daily energy consumption with noticeable variations on Thursdays, and much higher usage over the weekends. The other two households (1 and 4) tend to consume more energy with more  variation on the weekends relative to the week days, reflecting of work and leisure patterns.

@@

Figure \ref{fig:hod} shows energy consumption against time of day, separately by week day and weekend. Household 1 is an early bird, starting their day before 6 and going back home around 18 on week days. They switch air conditioning or heating on when they get home from work and keep it operating until mid-night, learned from the small horizontal cluster of points around 0.8 kWh. On the other hand, the stripes above 1 kWh for  household 2 indicates that perhaps air conditioning or heating runs continuously for some periods, consuming the twice the energy as household 1. A third peak occurs around 15 for household 3 only, likely when the kids are home from school. They also have a consistent energy pattern between week days and weekends. As for household 4, their home routine starts after 18 on week days. Figures \ref{fig:dow} and \ref{fig:hod}, part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.

@@

Figure \ref{fig:h1}-\ref{fig:h4} display the data in calendar layout individually for each household, unfolding their day-to-day life. Glancing over household 1, we can see that their overall energy use is low. Their week day energy use is distinguishable from their weekends, indicating a working household. The air conditioner appears to be used in the summer months (December--February) in the evening and weekends. In household 2, heating keeps functioning for consecutive hours, which is evident in the mid July. In contrast, household 1 uses heating cautiously. These observations help to explain the stripes and clusters in Figure \ref{fig:hod}. The calendar plots speak the stories about vacation time that are untold by previous plots. Household 1 is on vacation over three weeks of June, and household 2 was also away for vacation in late December and in the second week of June. Figure \ref{fig:h3} shows household 3 takes two one-month-long family trips in September until early October and in June/July. Household 4 is away over two or three weeks in early October, December, early April, and late June. The use of air conditioning and heating leaves no trace in these two households.

@@

# Discussion
\label{sec:discussion}

>>>>

bibliography: bibliography.bib
output: 
  rticles::asa_article:
    base_format: bookdown::pdf_document2
---

@@

The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped]. There have been 43 sensors installed that count pedestrians every hour across the inner-city area until the end of 2016 (Figure \@ref(fig:ped-map)). The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. The faceted plots in Figure \@ref(fig:time-series-plot) give an overall picture of the foot traffic at three different sensors over 2016. Further faceting by day of the week (Figure \@ref(fig:facet-time)) provides a better glimpse of the daily and sub-daily pedestrian patterns.

@@

```{r ped-map, fig.width = 7, out.width = "0.7\\linewidth", fig.cap = "(ref:ped-map-cap)"}
```

(ref:ped-map-cap) Map of the Melbourne city area with dots indicating sensor locations. These three highlighted sensors will be inspected in the paper: (1) the State Library---a public library, (2) Flagstaff Station---a train station, closed on non-work days, (3) Birrarung Marr---an outdoor park hosting many cultural and sports events.

```{r time-series-plot, fig.height = 6, fig.cap = "(ref:time-series-plot)"}
```

(ref:time-series-plot) Time series plots showing the number of pedestrians in 2016 measured at three different sensors in the city of Melbourne. Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number. A spike occurred at the State Library, caused by the annual White Night event on 20th of February. A relatively persistent pattern repeats from one week to another at Flagstaff Station. Birrarung Marr looks rather noisy and spiky, with a couple of chunks of missing records.

```{r facet-time, fig.height = 6, fig.cap = "(ref:facet-time-cap)"}
```

(ref:facet-time-cap) Hourly pedestrian counts for 2016 faceted by sensors and days of the week using lines. It primarily features two types of seasons---time of day and day of week---across all the sensors. Apparently other factors have influence over the number of pedestrians, which cannot be captured by the faceted plots, such as the overnight White Night traffic on Saturday at the State Library and a variety of events at Birrarung Marr.

@@

The remainder of the paper is organized as follows. Section \@ref(sec:algorithm) demonstrates the construction of the calendar layout in depth. Section \@ref(sec:transformation) describes the algorithms of data transformation. Section \@ref(sec:opt) lists and describes the options that come with the `frame_calendar` function. Section \@ref(sec:variations) presents some variations of its usage. Graphical analyses of sub-daily people's activities are illustrated with a case study in Section \@ref(sec:case). Section \@ref(sec:discussion) discusses the limitations of calendar displays and possible new directions.

# Creating a calendar display {#sec:algorithm}

## Data transformation {#sec:transformation}

@@

\begin{equation}
  \begin{aligned}
  i &= \lceil (g \text{ mod } 35) / 7\rceil, \\
  j &= g \text{ mod } 7. (\#eq:grid)
  \end{aligned}
\end{equation}

Figure \@ref(fig:month-diagram) illustrates this $(i,j)$ layout for a month where $k=5$.

```{r month-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "(ref:month-diagram-cap)"}
include_graphics("img/month.pdf")
```

(ref:month-diagram-cap) Illustration of the indexing layout for cells in a month, where $k$ is day of the week, $g$ is day of the month, $(i, j)$ indicates grid position.

To create the layout for a full year, $(m, n)$ denotes the position of the month arranged in the plot, where $1 \le m \le M$ and $1 \le n \le N$; $b$ denotes the small amount of white space between each month for visual separation. Figure \@ref(fig:year-diagram) illustrates this layout where $M = 3$ and $N = 4$.

```{r year-diagram, out.width = "360pt", out.height = "250pt", fig.cap = "(ref:year-diagram-cap)"}
include_graphics("img/year-diagram.pdf")
```

(ref:year-diagram-cap) Illustration of the indexing layout for months of one year, where $M$ and $N$ indicate number of rows and columns, $b$ is a space parameter separating cells.

@@

\begin{equation}
  \begin{aligned}
  x &= j + (n - 1) \times 7 + (n - 1) \times b + h, \\
  y &= i - (m - 1) \times 5 - (m - 1) \times b + c. (\#eq:final)
  \end{aligned}
\end{equation}

Note that for the vertical direction, the top left is the starting point of the grid (in Figure \@ref(fig:month-diagram)) which is why subtraction is performed. Within each cell, the starting position is the bottom left.

```{r fs-2016, fig.height = 7, fig.cap = "(ref:fs-2016-cap)"}
```

(ref:fs-2016-cap) The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs. The disparities between week day and weekend along with public holiday are immediately apparent. The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016. Note that the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs to May and October.

Figure \@ref(fig:fs-2016) shows the line glyphs framed in the monthly calendar over the year 2016. This is achieved by the `frame_calendar` function, which computes the coordinates on the calendar for the input data variables. These can then be plotted using the usual **ggplot2** R package [@R-ggplot2] functions. All of the grammar of graphics can be applied.

@@

Regarding the monthly calendar, the major reference lines separate every month panel and the minor ones separate every cell, represented by the thick and thin lines in Figure \@ref(fig:fs-2016), respectively. The major reference lines are placed surrounding every month block: for each $m$, the vertical lines are determined by $\min{(x)}$ and $\max{(x)}$; for each $n$, the horizontal lines are given by $\min{(y)}$ and $\max{(y)}$. The minor reference lines are only placed on the left side of every cell: for each $i$, the vertical division is $\min{(x)}$; for each $j$, the horizontal is $\min{(y)}$.

@@

## Options {#sec:opt}

@@

It is assumed that the `data` is in tidy format [@wickham2014tidy], and `x`, `y` are the variables that will be mapped to the horizontal and vertical axes in each cell. For example, the `x` is the time of the day, and `y` is the count (Figure \@ref(fig:fs-2016)). The `date` argument specifies the date variable used to construct the calendar layout.

@@

```{r fs-polar, fig.height = 7, fig.cap = "(ref:fs-polar-cap)", eval = FALSE}
```

(ref:fs-polar-cap) Figure \@ref(fig:fs-2016) in circular layout, which is referred to as star plots. The daily periodicity on work days are clearly visible.

@@

With local scaling, the overall variation gives way to the individual shape. Figure \@ref(fig:fs-free) shows the same data as Figure \@ref(fig:fs-2016) scaled locally using `scale = "free"`. The daily trends are magnified.

```{r fs-free, fig.height = 7, fig.cap = "(ref:fs-free-cap)"}
```

(ref:fs-free-cap) Line glyphs on the calendar format showing hourly foot traffic at Flagstaff Station, scaled over all the days. The individual shape on a single day becomes more distinctive, however it is impossible to compare the size of peaks between days.

@@

By default, grids are laid out horizontally. This can be transposed by setting the `dir` parameter to `"v"`, in which case $i$ and $j$ are swapped in Equation \@ref(eq:grid). This can be useful for creating  calendar layouts for countries where vertical layout is the convention.

@@

## Variations {#sec:variations}

@@

Plots can be layered. The comparison of sensors can be done by overlaying the values for each (Figure \@ref(fig:overlay)). Differences between the pedestrian patterns at these sensors can be seen. Flagstaff Station exhibits strong commuters patterns, with fewer pedestrian counts during the weekends and public holidays. This suggests that Flagstaff Station has limited functionality on non-work days. From Figure \@ref(fig:overlay) it can be seen that Birrarung Marr has a distinct temporal pattern from the other two all year round. The nighttime events, such as White Night, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to the State Library and Birrarung Marr.

```{r overlay, fig.height = 7, fig.cap = "(ref:overlay-cap)"}
```

(ref:overlay-cap) Overlaying line graphs of the three sensors in the monthly calendar. Three sensors demonstrate very different traffic patterns. Birrarung Marr tends to attract many pedestrians for special events held on weekends, contrasting to the bimodal commuting traffic at Flagstaff Station.

To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors. Figure \@ref(fig:facet) presents the idea of faceting calendar plots. This allows comparing the overall structure between sensors, while emphasizing individual sensor variation. In particular, it can be immediately learned that Birrarung Marr was busy and packed, for example during the Australian Open, a major international tennis tournament, in the last two weeks of January. This is concealed in the conventional graphics.

```{r facet, fig.height = 10.2, fig.cap = "(ref:facet-cap)"}
```

(ref:facet-cap) Line charts, embedded in the $6 \times 2$ monthly calendar, colored and faceted by the 3 sensors. The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors.

@@

The `frame_calendar` function is not constrained to line plots. The full range of plotting capabilities in **ggplot2** is essentially available. Figure \@ref(fig:scatterplot) shows a lag scatterplot at Flagstaff Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument. This figure is organized in the daily calendar layout. Figure \@ref(fig:scatterplot) indicates two primary patterns, strong autocorrelation on weekends, and weaker autocorrelation on work days. At the higher counts, on week days, the next hour sees possibly substantial increase or decrease in counts, essentially revealing a bimodal distribution of consecutive counts, as supported by Figure \@ref(fig:fs-2016).

```{r scatterplot, fig.height = 5, fig.cap = "(ref:scatterplot-cap)"}
```

(ref:scatterplot-cap) Lag scatterplot in the daily calendar layout. Each hour's count is plotted against previous hour's count at Flagstaff Station to demonstrate the autocorrelation at lag 1. The correlation between them is more consistent on non-work days than work days.

The algorithm can also produce more complicated plots, such as boxplots. Figure \@ref(fig:boxplot) uses a loess smooth line [@cleveland1979loess] superimposed on side-by-side boxplots. It shows the distribution of hourly counts across all 43 sensors during December. The last week of December is the holiday season: people are off work on the day before Christmas (December 24), go shopping on the Boxing day (December 26), and stay out for the fireworks on New Year's Eve.

@@

```{r chn, fig.height = 8, fig.cap = "(ref:chn-cap)", fig.showtext = TRUE}
```

(ref:chn-cap) The same plot as Figure \@ref(fig:boxplot), but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English.

```{r chn-embedded, fig.cap = "The same plot as Figure \\ref{fig:boxplot}, but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English.", eval = FALSE}
include_graphics("figure/chn-1.pdf")
```

@@

# Case study {#sec:case}

@@

Figure \@ref(fig:dow) shows the energy use across days of week in the form of letter value plots [@hofmann2017letter]. Letter value plots are a variant of boxplots for large data, with other quantiles represented by boxes. Letters indicate the fraction of the data divisions, for example, F indicates fourths or quartiles, and the two outer ends of the box are the 25th and 75th percentile, the same traditional ends of the box as a boxplot. The letter E indicates eighths, with box ends being 12.5th and 87.5th percentiles of the data. These additional boxes replace the whiskers in a traditional boxplot. The letter value plots for the households, show a line indicating the median (M) and the innermost boxes corresponding to the fourth (F) and the eighth (E) quantile divisions. Inspecting the medians across households tells us that household 3, a family size of one couple and two kids, uses more energy over the week days, than other households. The relatively larger boxes for household 2 indicates greater variability in daily energy consumption with noticeable variations on Thursdays, and much higher usage over the weekends. The other two households (1 and 4) tend to consume more energy with more  variation on the weekends relative to the week days, reflecting of work and leisure patterns.

@@

Figure \@ref(fig:hod) shows energy consumption against time of day, separately by week day and weekend. Household 1 is an early bird, starting their day before 6 and going back home around 18 on week days. They switch air conditioning or heating on when they get home from work and keep it operating until mid-night, learned from the small horizontal cluster of points around 0.8 kWh. On the other hand, the stripes above 1 kWh for  household 2 indicates that perhaps air conditioning or heating runs continuously for some periods, consuming the twice the energy as household 1. A third peak occurs around 15 for household 3 only, likely when the kids are home from school. They also have a consistent energy pattern between week days and weekends. As for household 4, their home routine starts after 18 on week days. Figures \@ref(fig:dow) and \@ref(fig:hod), part of a traditional graphical toolkit, are useful for summarizing overall deviations across days and households.

@@

Figure \@ref(fig:h1)-\@ref(fig:h4) display the data in calendar layout individually for each household, unfolding their day-to-day life. Glancing over household 1, we can see that their overall energy use is low. Their week day energy use is distinguishable from their weekends, indicating a working household. The air conditioner appears to be used in the summer months (December--February) in the evening and weekends. In household 2, heating keeps functioning for consecutive hours, which is evident in the mid July. In contrast, household 1 uses heating cautiously. These observations help to explain the stripes and clusters in Figure \@ref(fig:hod). The calendar plots speak the stories about vacation time that are untold by previous plots. Household 1 is on vacation over three weeks of June, and household 2 was also away for vacation in late December and in the second week of June. Figure \@ref(fig:h3) shows household 3 takes two one-month-long family trips in September until early October and in June/July. Household 4 is away over two or three weeks in early October, December, early April, and late June. The use of air conditioning and heating leaves no trace in these two households.

@@

# Discussion {#sec:discussion}

### 2019-03-12 main.tex
use base_foramt bookdown::pdf_document2

\hypertarget{creating-a-calendar-display}{%
\section{Creating a calendar
display}\label{creating-a-calendar-display}}

\label{sec:algorithm}

\hypertarget{data-transformation}{%
\subsection{Data transformation}\label{data-transformation}}

\label{sec:transformation}

@@

\caption{Illustration of the indexing layout for cells in a month, where $k$ is day of the week, $g$ is day of the month, $(i, j)$ indicates grid position.}\label{fig:month-diagram}
\end{figure}

@@

\caption{Illustration of the indexing layout for months of one year, where $M$ and $N$ indicate number of rows and columns, $b$ is a space parameter separating cells.}\label{fig:year-diagram}
\end{figure}

@@

\caption{The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs. The disparities between week day and weekend along with public holiday are immediately apparent. The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016. Note that the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs to May and October.}\label{fig:fs-2016}
\end{figure}

@@

\hypertarget{options}{%
\subsection{Options}\label{options}}

\label{sec:opt}

@@

By default, grids are laid out horizontally. This can be transposed by
setting the \texttt{dir} parameter to \texttt{"v"}, in which case \(i\)
and \(j\) are swapped in Equation \ref{eq:grid}. This can be useful for
creating calendar layouts for countries where vertical layout is the
convention.

@@

\hypertarget{variations}{%
\subsection{Variations}\label{variations}}

\label{sec:variations}

@@

\caption{Line charts, embedded in the $6 \times 2$ monthly calendar, colored and faceted by the 3 sensors. The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors.}\label{fig:facet}
\end{figure}

@@

\caption{The same plot as Figure \ref{fig:boxplot}, but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English.}\label{fig:chn-embedded}
\end{figure}

@@

\hypertarget{case-study}{%
\section{Case study}\label{case-study}}

\label{sec:case}

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\label{sec:discussion}

>>>>

\hypertarget{sec:algorithm}{%
\section{Creating a calendar display}\label{sec:algorithm}}

\hypertarget{sec:transformation}{%
\subsection{Data transformation}\label{sec:transformation}}

@@

\caption{Illustration of the indexing layout for cells in a month, where \(k\) is day of the week, \(g\) is day of the month, \((i, j)\) indicates grid position.}\label{fig:month-diagram}
\end{figure}

@@

\caption{Illustration of the indexing layout for months of one year, where \(M\) and \(N\) indicate number of rows and columns, \(b\) is a space parameter separating cells.}\label{fig:year-diagram}
\end{figure}

@@

\caption{The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs. The disparities between week day and weekend along with public holiday are immediately apparent. The arrangement of the data into a \(3 \times 4\) monthly grid represents all the traffic in 2016. Note that the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs to May and October.}\label{fig:fs-2016}
\end{figure}

@@

\hypertarget{sec:opt}{%
\subsection{Options}\label{sec:opt}}

@@

By default, grids are laid out horizontally. This can be transposed by setting the \texttt{dir} parameter to \texttt{"v"}, in which case \(i\) and \(j\) are swapped in Equation \eqref{eq:grid}. This can be useful for creating calendar layouts for countries where vertical layout is the convention.

@@

\hypertarget{sec:variations}{%
\subsection{Variations}\label{sec:variations}}

@@

\caption{Line charts, embedded in the \(6 \times 2\) monthly calendar, colored and faceted by the 3 sensors. The variations of an individual sensor are emphasised, and the shapes can be compared across the cells and sensors.}\label{fig:facet}
\end{figure}

@@

\caption{The same plot as Figure \ref{fig:boxplot}, but with the month and week day labels in Chinese. It demonstrates the natural support for languages other than English.}\label{fig:chn}
\end{figure}

@@

\hypertarget{sec:case}{%
\section{Case study}\label{sec:case}}

@@

\hypertarget{sec:discussion}{%
\section{Discussion}\label{sec:discussion}}
